\section{Broader ethical concerns}
\label{sec:broader}

A number of broader ethical concerns relating to the project were also identified:
\begin{itemize} 
\item
\textbf{Dual use}. There is clear potential for the use of \project technologies for military and defence purposes.  Although the project is not pursuing such applications, we should maintain awareness of such risks.  It is a fact that the development of \project tools as open source both increases the probability of socially beneficial uses of \project technology and their use for defence or intelligence gathering.  
\item
\textbf{Job security and automation}. \project will develop tools that will have impact on people's existing jobs.  As a project we should consider whether \project technology will replace jobs, deskill existing jobs, or enrich existing jobs.  How will \project technology affect the roles of experts~--~will they still be required?   Will it increase their capacity?
\item 
\textbf{Data journalism}.  It is important to perform some kind of scenario modelling related to the data journalism use case that is being developed.  How will journalists use \project tools after the completion of the project?  What will be the effect of \project technologies on journalism over the next decade, as automatic processing of content and machine learning become ubiquitous and mature technologies?  One  particular issue is how to deal with the biases that will arise from automatic distillation of large-scale news content. How do we frame and address this challenge?
\end{itemize}
