\section{Import and export of data to non-EU countries}
\label{sec:nonEU}

Import and export of data to and from non-EU countries raises legal and ethical issues which need to be addressed.

The project has participation from Qatar (QCRI) and Switzerland (IDIAP).  Although Switzerland is not a member of the EU, the Swiss Federal Data Protection Act (19 June 1992) follows similar concepts to the EU Data Protection Directive. Accordingly, the European Commission has found Switzerland to provide an adequate level of data protection from an EU perspective (Decision 2000/518/EC). 

QCRI are a Linked Third Party (to UEDIN). This raises the legal issue of the export of personal data outside the EU.  In the \project project, this is most important to social media data.  Since this data is only used in WP5 -- Task T5.3, to which QCRI does not participate, it is  straightforward to comply with this legal requirement.

A second issue relates to the processing of broadcast content from Ukraine, Iran, Russia, and a number of Arabic countries.  The project partners' use of this data fall under the following annex to the consortium agreement:
\begin{quote}
{\it 
Unless agreed otherwise in a separate BBC licence, any Access Rights to datasets in accordance with the Agreement, shall be provided by the BBC (i) as a sample from the operational sources of the BBC (including BBC Monitoring and the operational production systems of BBC World Service); and (ii) only for implementation (where a copy of the dataset or parts thereof are provided, such copy shall be returned to the BBC or deleted in accordance with any other conditions that the BBC may apply).

BBC licences to third-party datasets (including third-party news and programming content) cannot be passed on via licensing and are expressly excluded from exploitation.}
\end{quote}


    
    